VM-Centric Snapshot Deduplication with Fault Isolation for Converged Cloud Architectures

### 2014-01-30 batchdedup/intro.tex

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host
%tens of thousands of  virtual machines(VM) created for various active users.
Periodic  archiving of virtual machine (VM) snapshots is important 
for long-term data  retention and fault recovery. 
%in case of machine failures.  
For example, daily backup of VM images  is conducted automatically 
at Alibaba which provides the largest public cloud service in China.
The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
This issue has been addressed by   storage data deduplication ~\cite{venti02,bottleneck08} that
identifies redundant content duplicates among snapshots.  
One architectural approach is to attach  a separate backup system with deduplication
support to the cloud cluster, and  every machine periodically transfers snapshots to
the attached backup system.  Such a dedicated backup  configuration can be expensive, 
considering that significant networking  and computing resource  is required to transfer 
raw data and conduct signature comparison.

This paper seeks for a low-cost architecture option and considers a backup
service that uses the existing cloud computing resource.
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized. 
Performing deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  small or modest resources 
%with a minimal impact to the existing cloud services.  Another challenge for backup with deduplication is 
with a minimal impact to the existing cloud services.  Another challenge is 
%that deletion of old snapshots compete for computing resource as well. That is because data dependence created 
that deletion of old snapshots compete for computing resource as well, because data dependence created 
by duplicate relationship among snapshots  adds processing complexity.
% especially when  VMs can migrate around in the cloud.

>>>>

%Virtualization is the engine behind many popular cloud computing platforms.
%Amazon, Alibaba,  and many others have provided public VM clouds that host
%tens of thousands of  virtual machines(VM) created for various active users.
Periodic  archiving of virtual machine (VM) snapshots is important 
for long-term data  retention and fault recovery. 
%in case of machine failures.  
For example, daily backup of VM images  is conducted automatically 
at Alibaba which provides the largest public cloud service in China.
The cost of frequent backup of VM snapshots is  high because of the huge storage demand.
This issue has been addressed by   storage data deduplication ~\cite{venti02,bottleneck08} that
identifies redundant content duplicates among snapshots.  
One architectural approach is to attach  a separate backup system with deduplication support such as ones
from  EMC~\cite{EMC} or NetApp{NetApp}
to the cloud cluster, and  the snaphosts of every virtual machine is   periodically transferred to
the attached backup system.  Such a dedicated backup  configuration can be expensive, 
considering that significant networking  and computing resource  is required 
to transfer undeduplicatd data and conduct signature comparison. 
Our work is targeted applications where data processing and archiving  can be conducted
in an inexpensive commodity storage cluster using the  Google file system or Hadoop~\cite{googlefs03}
and such a storage cluster is non-dedicated, shared with other services running in the entire cluster.
Resource sharing is common especially in the cloud where many computing and storage
services are collocated in the same physical machines.

With  resource sharing with other services as an assumption, we are motivated to develop a 
low-profile backup solution with deduplication. 
Deduplication is a major

%This paper seeks for a low-cost architecture option and considers a backup
%service that uses the existing cloud computing resource.
%Since most of backup data is not used in practice, system resource in such a service is not fully utilized. 
Performing deduplication adds significant  memory cost for comparison of content fingerprints. 
Since each physical machine in a cluster  hosts many VMs, memory contention happens frequently. 
Cloud providers often wish that the backup service only consumes  small or modest resources 
%with a minimal impact to the existing cloud services.  Another challenge for backup with deduplication is 
with a minimal impact to the existing cloud services.  Another challenge is 
%that deletion of old snapshots compete for computing resource as well. That is because data dependence created 
that deletion of old snapshots compete for computing resource as well, because data dependence created 
by duplicate relationship among snapshots  adds processing complexity.
% especially when  VMs can migrate around in the cloud.

### 2014-01-30 snapshot_system/paper_dedup.tex

Since $k$ is relatively small and these top $k$ chunks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular chunk data to enhance the fault resilience.

Figure~\ref{fig:Datazipf} illustrates the distribution of chunk popularity based on a data trace from Alibaba with over 2500 VMs.
The distribution is zipf like and popular chunks are dominating the large portion of data chunks.
We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
$\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.

@@

\begin{figure}
\centering
%\epsfig{file=log-log.disk,width=3in}
%source ../vm_snapshot_sim/images/log-log.disk.eps
 \includegraphics[width=3in]{figures/log-log-disk}
\caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
\label{fig:Datazipf}
\end{figure}

>>>>

Since $k$ is relatively small and these top $k$ chunks are shared among multiple VMs, 
we can afford to provide extra replicas for these popular chunk data to enhance the fault resilience.
We have studied the distribution of chunk popularity~\cite{WeiZhangIEEE} 
based on a data trace from Alibaba with over 2500 VMs.
The distribution is zipf-like and popular chunks are dominating the large portion of data chunks.
We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
$\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.

% Figure~\ref{fig:Datazipf} illustrates the distribution of chunk popularity based on a data trace from Alibaba with over 2500 VMs.
% The distribution is zipf like and popular chunks are dominating the large portion of data chunks.
% We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
% $\sigma$ with a range of 1 to 4\% can deliver a fairly competitive deduplication efficiency.

@@

% \begin{figure}
% \centering
%  \includegraphics[width=3in]{figures/log-log-disk}
% \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
% \label{fig:Datazipf}
% \end{figure}
